---
title: "Fairness in Winner-Take-All Markets: replication package"
author: "Christian Braathen and Erik O. Sorensen."
date: "May 1 2018"
output: html_document
---

The code in this file is developed in parallell with the code
used to estimate the results in the paper, and also by a different 
team in order to ensure that reported analysis and estimation code
matches up.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(multcomp)
library(tidyverse)
library(latex2exp)    # Annotate plots with LaTeX
library(cowplot)      # Arranging plots in a grid
library(stargazer)    # Formatting regression tables
library(multiwayvcov) # Robust covariance matrix
```


# Reading data and initial data transformations.
Reading inn data from file. 
```{r load_data}
df_data <- read.csv("data/mmwinner.csv",sep=",")
```

First, make the factor with treatments with names that correspond to those
in the paper
```{r data_transformation_1}
df_data$T <- as.factor(df_data$T)
levels(df_data$T) <- c("WTA-No Choice",
                       "WTA",
                       "WTA-No Exp.",
                       "Base")
```

Some transformations that are useful for producing the graphs.
```{r data_transformation_2}
data_no_ties  <- df_data %>% mutate( shareWinner = 1 - shareX,
                                     allWinner = (e2==y2),
                                     difference = abs(x1-x2),
                                     performance_winner = x2,
                                     performance_loser = x1,
                                     normalized_income_winner = y2/0.15,
                                     normalized_income_loser  = y1/0.15,
                                     difference_trimmed = ifelse(difference>10,10,difference),
                                     cright = political>=4,
                                     female = (sex==2)) %>%
  filter(e1==0) 
```


# Figure 1
Figure 1 plots a histogram of the share given to the winner. We plot in 1/11 wide
bins so that the center bin is centered at 1/2.

We want to reorder the treatments for this purpose.
```{r, echo=FALSE}
df_figure1 <- data_no_ties %>% mutate(T = factor(T,levels(T)[c(2,4,1,3)]))
ggplot(df_figure1) +
  geom_histogram(aes(shareWinner, (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
           colour="black",
           fill="#808080",
           boundary = 0,
           binwidth= 0.09090909090909091) +
  facet_wrap(~(T),
             scales="fixed") +
  scale_x_continuous(name="Share given to winner",
                     breaks=c(0,0.5, 1)) +
  ylab("Fraction") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

# Figure 2 
Figure 2 plots the share to winner and the mean share giving all to the winner by
the winning margin.
First, summarizing the data to plotted. Note that the top we are summarizing
by a trimmed difference.
```{r, echo=FALSE}
data_no_ties_figure_2 <- data_no_ties %>%
  filter(T!="Base") %>%
  group_by(difference_trimmed) %>%
  summarize( mean_shareWinner = mean(shareWinner),
             se_shareWinner = sd(shareWinner)/sqrt(n()),
             mean_allWinner = mean( y2==e2),
             se_allWinner = sd( y2==e2)/sqrt(n()))
```

First the share given to the winner:
```{r}
plot_shareToWinner_pooled <- data_no_ties_figure_2 %>% 
  ggplot(aes(x = difference_trimmed, y = mean_shareWinner)) +
  geom_errorbar(aes(ymin=mean_shareWinner - se_shareWinner, 
                    ymax=mean_shareWinner + se_shareWinner),
                colour="#A0A0A0", width=0.1) +
  geom_line() +
  geom_point(colour="#606060") +
  scale_x_continuous(name="Performance difference",
                     breaks = round(seq(min(data_no_ties_figure_2$difference_trimmed), 
                                        max(data_no_ties_figure_2$difference_trimmed), 
                                        by = 1),0),
                     labels=c(1:9,TeX('$\\geq 10$'))) +
  scale_y_continuous(name="Mean share given to winner",
                     breaks = seq(0.5, 1, by = 0.1),
                     limits = c(0.5,1)) +
  theme_bw() +
  ggtitle("Share to Winner") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))

```

Now, similar, but for the share that gives everything to the winner:

```{r, echo=FALSE}
plot_allToWinner_pooled <- data_no_ties_figure_2 %>% 
  ggplot(aes( x = difference_trimmed, y= mean_allWinner)) + 
  geom_errorbar( aes( ymin= mean_allWinner - se_allWinner,
                      ymax = mean_allWinner + se_allWinner),
                 colour="#A0A0A0", width=0.1) +
  geom_line() + 
  geom_point(colour="#606060") +
  scale_x_continuous(name="Performance difference",
                     breaks = round(seq(min(data_no_ties_figure_2$difference_trimmed), 
                                        max(data_no_ties_figure_2$difference_trimmed), 
                                        by = 1),0),
                     labels=c(1:9,TeX('$\\geq 10$'))) +  
  scale_y_continuous(name="Mean share giving all to winner",
                     breaks = seq(0, 1, by = 0.2),
                     limits = c(0,1)) +
  theme_bw() +
  ggtitle("All to Winner") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))
```
Now for putting these two panels together:

```{r, echo=FALSE}
plot_grid(plot_shareToWinner_pooled, 
          plot_allToWinner_pooled,
          ncol=1)
```

# Figure 3
In figure 3, we want to plot mean incomes of winner against winner performance 
and mean income of loser against the performance of the loser, with lines for indicating
the hypotheses of factual merit (left) and counterfactual merit (right).

First, preparing the data
```{r, echo=FALSE, warning=FALSE}
F3L <- data_no_ties %>%
  filter(T!="Base") %>%
  group_by(performance_winner) %>%
  summarize( mean_y = mean(normalized_income_winner),
             se_y = sd(normalized_income_winner)/sqrt(n()))
F3R <- data_no_ties %>%
  filter(T!="Base") %>%
  group_by(performance_loser) %>%
  summarize(mean_y = mean(normalized_income_loser),
            se_y = sd(normalized_income_loser)/sqrt(n()))
```

Now, we want to plot the left hand side figure
```{r}
plot_income_winner <- F3L %>% 
  ggplot(aes(x=performance_winner, y=mean_y)) + 
  geom_errorbar(aes(ymin=mean_y-se_y, ymax=mean_y +se_y),
                colour="#808080", 
                width=0.6,
                size=0.1) +
  geom_line() +
  geom_abline(aes(intercept=0,slope=1)) +
  geom_point(colour="#606060") +
  scale_x_continuous(name="Performance of winner",
                     breaks = seq(0, 24, by = 4),
                     limits=c(0,24)) +
  scale_y_continuous(name=TeX('Mean income of winner \\[normalized\\] $\\pm$ s.e.m.'),
                     breaks = seq(0, 24, by = 4),
                     limits=c(0,24)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  theme(aspect.ratio=1)
```

The right hand side in similar way:
```{r}
plot_income_loser <- F3R %>% 
  ggplot(aes(x=performance_loser, y=mean_y)) + 
  geom_errorbar(aes(ymin=mean_y-se_y, ymax=mean_y+se_y),
                colour="#808080", 
                width=0.6,
                size=0.1) +
  geom_line() +
  geom_abline(aes(intercept=0,slope=0.5)) +
  geom_point(colour="#606060") +
  scale_x_continuous(name="Performance of loser",
                     breaks = seq(0, 24, by = 4),
                     limits=c(0,24)) +
  scale_y_continuous(name=TeX('Mean income of loser \\[normalized\\] $\\pm$ s.e.m.'),
                     breaks = seq(0, 24, by = 4),
                     limits=c(0,24)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  theme(aspect.ratio=1)
```
Now putting them together in a single plot

```{r, echo=FALSE, warning=FALSE}
plot_grid(plot_income_winner, 
          plot_income_loser,
          ncol=2)
```


# Regression tables


## Regression results: Treatment effects (Table 2)
The regression table in the main text looks at treatment effects on 
the share to the winner and the share that gives all to the winner. 
I create the treatment dummies explicitly to control the names in 
the regression output.

```{r treatment_regressions}
dR1 <- data_no_ties %>%
  mutate(wta = T=="WTA",
         wtanc = T=="WTA-No Choice",
         wtane = T=="WTA-No Exp.")
lmR1_1 <- lm(shareWinner ~ wta + wtanc + wtane, data = dR1)
vR1_1 <- cluster.vcov(lmR1_1, 1:nrow(dR1))
lmR1_2 <- lm(shareWinner ~ wta + wtanc + wtane + age + female + cright + I(sessionid), 
             data = dR1)
vR1_2 <- cluster.vcov(lmR1_2, 1:nrow(dR1))
lmR1_3 <- lm(allWinner ~ wta + wtanc + wtane , data = dR1)
vR1_3 <- cluster.vcov(lmR1_3, 1:nrow(dR1))
lmR1_4 <- lm(allWinner ~ wta + wtanc + wtane + age + female + cright + I(sessionid), 
             data = dR1)
vR1_4 <- cluster.vcov(lmR1_4, 1:nrow(dR1))
```

Calculation of some ancillary output. 
```{r}
# First, dependent mean in Base:
m1 <- sprintf("%4.3f", mean(dR1$shareWinner[dR1$T=="Base"]))
m2 <- sprintf("%4.3f", mean(dR1$allWinner[dR1$T=="Base"]))
R1_means <- c("Dependent mean (base):", m1,m1,m2,m2)
# Three rows of Linear hypothesis tests
p1 <- c("P: WTA = WTA-No choice",
        sprintf("%4.3f", summary(glht(lmR1_1, linfct=c("wtaTRUE - wtancTRUE = 0"),
                                     vcov=vR1_1))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_2, linfct=c("wtaTRUE - wtancTRUE = 0"),
                                      vcov=vR1_2))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_3, linfct=c("wtaTRUE - wtancTRUE = 0"),
                                      vcov=vR1_3))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_4, linfct=c("wtaTRUE - wtancTRUE = 0"),
                                      vcov=vR1_4))$test$pvalues[1]))
p2 <- c("P: WTA = WTA-No Exp.",
        sprintf("%4.3f", summary(glht(lmR1_1, linfct=c("wtaTRUE - wtaneTRUE = 0"),
                                     vcov=vR1_1))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_2, linfct=c("wtaTRUE - wtaneTRUE = 0"),
                                      vcov=vR1_2))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_3, linfct=c("wtaTRUE - wtaneTRUE = 0"),
                                      vcov=vR1_3))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_4, linfct=c("wtaTRUE - wtaneTRUE = 0"),
                                      vcov=vR1_4))$test$pvalues[1]))
p3 <- c("P: WTA No choice = WTA-No Exp.",
        sprintf("%4.3f", summary(glht(lmR1_1, linfct=c("wtaneTRUE - wtancTRUE = 0"),
                                     vcov=vR1_1))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_2, linfct=c("wtaneTRUE - wtancTRUE = 0"),
                                      vcov=vR1_2))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_3, linfct=c("wtaneTRUE - wtancTRUE = 0"),
                                      vcov=vR1_3))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmR1_4, linfct=c("wtaneTRUE - wtancTRUE = 0"),
                                      vcov=vR1_4))$test$pvalues[1]))
```

Output of regressions into table
```{r}
stargazer( lmR1_1, lmR1_2, lmR1_3, lmR1_4,
           se =list(sqrt(diag(vR1_1)), sqrt(diag(vR1_2)), 
                    sqrt(diag(vR1_3)), sqrt(diag(vR1_4))),
           type="text", 
           column.labels=c("Share to winner", "All to winner"),
           column.separate=c(2,2),
           df=FALSE, 
           omit = c("I(sessionid)*"),
           keep.stat=c("adj.rsq", "n"),
           add.lines = list(c("Session FE","no","yes","no","yes"),
                              R1_means, p1, p2, p3)
)
```


## Regression results: Counterfactual vs Factual merit (Table A.2)
```{r factual_counterfactual_regressions}
dRA2 <- data_no_ties %>% filter(T!="Base") %>%
  rename( performance_winner = x2, performance_loser = x1)
lmRA2_1 <- lm(normalized_income_winner ~ performance_winner, data = dRA2)
vRA2_1 <- cluster.vcov(lmRA2_1, 1:nrow(dRA2))
lmRA2_2 <- lm(normalized_income_winner ~ performance_winner + 
                age + female + cright + I(sessionid), data = dRA2)
vRA2_2 <- cluster.vcov(lmRA2_2, 1:nrow(dRA2))
lmRA2_3 <- lm(normalized_income_loser ~ performance_loser, data = dRA2)
vRA2_3 <- cluster.vcov(lmRA2_3, 1:nrow(dRA2))
lmRA2_4 <- lm(normalized_income_loser  ~ performance_loser + 
                age + female + cright + I(sessionid), data = dRA2)
vRA2_4 <- cluster.vcov(lmRA2_4, 1:nrow(dRA2))
```

Some ancillary tests are needed:
```{r}
pA2_1<- c("P: Performance winner = 1", 
        sprintf("%4.3f", summary(glht(lmRA2_1, linfct=c("performance_winner = 1"),
                                     vcov=vRA2_1))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmRA2_2, linfct=c("performance_winner = 1"),
                                     vcov=vRA2_2))$test$pvalues[1]),
        " ", " ")

pA2_2<- c("P: Performance loser = 0.5", " ", " ",
        sprintf("%4.3f", summary(glht(lmRA2_3, linfct=c("performance_loser = 0.5"),
                                     vcov=vRA2_3))$test$pvalues[1]),
        sprintf("%4.3f", summary(glht(lmRA2_4, linfct=c("performance_loser = 0.5"),
                                     vcov=vRA2_4))$test$pvalues[1]))

```

Output of regressions into table
```{r}
stargazer( lmRA2_1, lmRA2_2, lmRA2_3, lmRA2_4,
           se =list(sqrt(diag(vRA2_1)), sqrt(diag(vRA2_2)), sqrt(diag(vRA2_3)), sqrt(diag(vRA2_4))),
           type="text",
           column.labels = c("Income to winner", "Income to loser"),
           column.separate=c(2,2),
           df=FALSE, 
           order=c("performance_winner", "performance_loser"),
           keep.stat=c("adj.rsq", "n"),
           omit = c("I(sessionid)*"),
          add.lines=list(c("Session FE", "no", "yes", "no", "yes"),
                         pA2_1, pA2_2))
```

# Further appendix material

## Randomization into treatments
First the summaries of data:
```{r}
data_no_ties %>% group_by(T) %>%
  summarize(mean_age = mean(age),
            mean_female = mean(female),
            mean_right = mean(cright))
data_no_ties %>% summarize(mean_age = mean(age),
            mean_female = mean(female),
            mean_right = mean(cright))
```

Now, Anova and chi square tests by variable as relevant:
```{r}
summary(aov(age ~ T, data = data_no_ties))
chisq.test(table(data_no_ties$female, data_no_ties$T))
chisq.test(table(data_no_ties$cright, data_no_ties$T))
```

## Histogram of performance by treatment (Figure A.1)
For the histogram of performance, it is easiest to add a an "All" group
and then produce the 4-panel graph with ggplot's facet feature.
```{r}
data_no_ties %>% 
  filter(T!="Base") %>%
  mutate(T = factor("All WTA treatments")) %>% 
  bind_rows(data_no_ties) %>%
  filter(T!="Base") %>%
  mutate(treatment = factor(T, levels=c("All WTA treatments",
                                        "WTA", 
                                        "WTA-No Choice",
                                        "WTA-No Exp."))) %>%
  gather(status, x, x1:x2) %>%
  ggplot(aes(x=x, y= (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
    geom_histogram(binwidth=1, boundary=0, colour="black",fill="#808080") + 
    theme_bw() + facet_wrap(~treatment) + ylab("Fraction") + xlab("Performance of worker")
```

## Histogram of performance difference by treatment (Figure A.2)
Almost same procedure as above
```{r}
data_no_ties %>% 
  filter(T!="Base") %>%
  mutate(T = factor("All WTA treatments")) %>% 
  bind_rows(data_no_ties) %>%
  filter(T!="Base") %>%
  mutate(treatment = factor(T, levels=c("All WTA treatments",
                                        "WTA", 
                                        "WTA-No Choice",
                                        "WTA-No Exp.")),
         performance_difference = abs(x2-x1))  %>%
  ggplot(aes(x = performance_difference, 
             y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..] )) +
    geom_histogram(binwidth=1, colour="black",fill="#808080") +
    theme_bw() + facet_wrap(~treatment) + ylab("Fraction") + xlab("Performance difference")
```

## The winning margin by treatment

First, creating data
```{r}
data_no_ties_figure_A3_A4 <- data_no_ties %>%
  filter(T!="Base") %>%
  group_by(T, difference_trimmed) %>%
  summarize( mean_shareWinner = mean(shareWinner),
             se_shareWinner = sd(shareWinner)/sqrt(n()),
             mean_allWinner = mean( y2==e2),
             se_allWinner = sd( y2==e2)/sqrt(n())) %>%
  mutate( treatment = fct_relevel(T, c("WTA", "WTA-No Choice", "WTA-No Exp.")))
```

### Figure A3:
```{r}
data_no_ties_figure_A3_A4 %>% 
  ggplot(aes(x = difference_trimmed, y = mean_shareWinner)) +
  geom_errorbar(aes(ymin=mean_shareWinner - se_shareWinner, 
                    ymax=mean_shareWinner + se_shareWinner),
                colour="#A0A0A0", width=0.1) +
  geom_line() +
  geom_point(colour="#606060") +
  scale_x_continuous(name="Performance difference",
                     breaks = round(seq(min(data_no_ties_figure_A3_A4$difference_trimmed), 
                                        max(data_no_ties_figure_A3_A4$difference_trimmed), 
                                        by = 1),0),
                     labels=c(1:9,TeX('$\\geq 10$'))) +
  scale_y_continuous(name="Mean share given to winner",
                     breaks = seq(0.5, 1, by = 0.1),
                     limits = c(0.5,1)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) + facet_wrap(~treatment, ncol=1)

```

### Figure A4: 
```{r}
data_no_ties_figure_A3_A4 %>% 
  ggplot(aes(x = difference_trimmed, y = mean_allWinner)) +
  geom_errorbar(aes(ymin=mean_allWinner - se_allWinner, 
                    ymax=mean_allWinner + se_allWinner),
                colour="#A0A0A0", width=0.1) +
  geom_line() +
  geom_point(colour="#606060") +
  scale_x_continuous(name="Performance difference",
                     breaks = round(seq(min(data_no_ties_figure_A3_A4$difference_trimmed), 
                                        max(data_no_ties_figure_A3_A4$difference_trimmed), 
                                        by = 1),0),
                     labels=c(1:9,TeX('$\\geq 10$'))) +
  scale_y_continuous(name="Mean share giving all to winner",
                     breaks = seq(0.0, 1, by = 0.2),
                     limits = c(0.0,1)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) + facet_wrap(~treatment, ncol=1)
```

# Versions of packages used
```{r}
sessionInfo()
```

